---
title: "Statistical analysis of tissue-specific miRNA and gene expression"
author: "Yining Ding"
output:
  html_notebook:
    toc: yes
    toc_depth: 3
bibliography: iid.bib
---
# Libraries
* The libraries used in this notebooks are
  * ggrepel [@ggrepel]
  * forcats [@forcats]
  * scales [@scales]
  * httr [@httr]
  * miRNAmeConverter [@miRNAmeConverter]
  * dplyr [@dplyr]
  * ggplot2 [@ggplot2]
  * HGNChelper [@HGNChelper]
  * kableExtra [@kableExtra]
  * biomaRt [@biomaRt]
  * tidyverse [@tidyverse]
  * ggpubr [@ggpubr]
```{r, echo=FALSE}
library(dplyr)
library(ggplot2)
library(ggrepel)
library(forcats)
library(scales)
library(ggpubr)
library(R.utils)
```

```{r child='mir_expr_table.Rmd', include=TRUE, eval=TRUE}
```

# Query genes associated with my miR
* Here we used the tables that combined all the expression miRNAs and query for their corresponding gnes. We found that the number of unique genes are 15536.
```{r, results='hide'}
combined_all_expr_tables <- read.table("/Users/helending/Documents/BCB330/mir_expr_table/data/combined_all_expr_tables.csv")

# query for their corresponding genes
mirs <- paste(combined_all_expr_tables$miR, collapse = ", ")
source("/Users/helending/Documents/BCB330/mir_expr_table/api.R")
mirDIP_genes <- get_mirDIP_gene(microrna_string = mirs, 
                                userScore = "Very High")
unique_mirdip <- unique(mirDIP_genes$Gene.Symbol)# length: 15536
```

```{r child='/Users/helending/Documents/BCB330/gtex/gtex.Rmd', include=TRUE, eval=TRUE}
```

# IID
## gene-gene interaction
* Get IID human annotated PPIs [@kotlyar2019iid] genes, convert the two column gene-gene interactions into one column by copying and pasting one column to the bottom of the other column. And then we removed duplicates and write to the file: __gene_interaction_dup.csv__, and saved in the __data__ folder.
```{r API, results='hide'}

IID_data <- read.delim("/Users/helending/Documents/BCB330/mir_expr_table/data/human_annotated_PPIs.txt", header = TRUE, sep = "\t")
gene_interaction <- IID_data
nrow(gene_interaction) # 975877

gene_interaction_copy <- gene_interaction
gene_interaction_copy[,c(1:4)] <- gene_interaction_copy[, c(1, 2, 4, 3)]

# bind the same table under the existing one, flip, and delete the "symbol 2" column
gene_interaction <- rbind(gene_interaction, gene_interaction_copy)

gene_interaction <- gene_interaction[,-c(1, 2, 4)]
colnames(gene_interaction)[1] <- "SYMBOLS"
gene_interaction <- gene_interaction[, -c(2, 3, 4, 5)]

# Remove duplicates
duplicated_genes <- which(duplicated(gene_interaction[1])) # 18726 unique genes
gene_interaction_dup <- gene_interaction
gene_interaction_dup <- gene_interaction_dup[-duplicated_genes,]
nrow(gene_interaction_dup) # 18726

write.table(gene_interaction_dup, "/Users/helending/Documents/BCB330/mir_expr_table/data/gene_interaction_dup.csv")
```

## Corresponding genes and tissues in IID
* Here we retrieve corresponding genes and corresponding tissues (15 tissues that we have data for miR expression data).
```{r}
copy_y <- read.table("/Users/helending/Documents/BCB330/mir_expr_table/data/gene_interation_dup.csv", stringsAsFactors = FALSE)

# Only examine the 15 tissues
tissue_names <- character()
for (i in 1:length(new_tables_0.99)){
  tissue_names[i] <- colnames(new_tables_0.99[[i]])[2]
}
copy_y_copy <- copy_y[which(copy_y$SYMBOLS %in% unique_mirdip), which(colnames(copy_y) %in% tissue_names)]
```

## IID pie graph
* Here we generates an overview of the IID annotated data. 
```{r}
unknown_sum <- length(copy_y_copy[copy_y_copy == "?"]) # 39
has_sum <- length(copy_y_copy[copy_y_copy == "1"]) # 489
no_sum <- length(copy_y_copy[copy_y_copy == "0"]) # 327
df_iid <- data.frame(
  group = c("?", "1", "0"),
  value = c(unknown_sum, has_sum, no_sum))
df_iid %>%
arrange(desc(value)) %>%
mutate(prop = percent(value / sum(value))) -> mydf 

ggplot(mydf, aes(x = "", y = value, fill = fct_inorder(group))) +
       geom_bar(width = 1, stat = "identity") +
       coord_polar("y", start = 0) +
       geom_label_repel(aes(label = prop), size=5, show.legend = F, nudge_x = 1) +
       guides(fill = guide_legend(title = "Group"))

copy_y_copy <- cbind("gene" = copy_y$SYMBOLS[which(copy_y$SYMBOLS %in% unique_mirdip)], copy_y_copy)
```

## IID statistics (number 1)
* Here we examine the general statistics of IID gene expression data by adding a column and sum rowise. The biological meaning is that we want to determine whether gene in IID are tissue specific or not by examining the number of tissues a miRNA targets.
```{r}
# add a column
copy_y_copy <- cbind(copy_y_copy, "num_tissue" = rowSums(copy_y_copy == "1"))
num_tissue_table <- data.frame(table(copy_y_copy$num_tissue))
colnames(num_tissue_table) <- c("num_tissues", "num_gene")


num_tissue_table_percent <- num_tissue_table
num_tissue_table_percent$per <- paste0(round(num_tissue_table_percent$num_gene / sum(num_tissue_table_percent$num_gene)*100, digits = 1), '%')
rownames(num_tissue_table_percent) <- 1:nrow(num_tissue_table_percent)

# graph
ggplot(num_tissue_table_percent, aes(x = num_tissues, y = num_gene))+
  geom_text(aes(label = per), size = 3, hjust = 0.5, vjust = -1, position = "stack")+
  geom_col(width = 0.8)
```

## IID statistics (number 2)
* Here we are at the second stage of examining the overal statistics by adding a row. The biological meaning is that we want to know for each tissue, what is the number of miRNA that targets it. Therefore, getting the sense that whether the tissue contains tissue specific or generic miRNA expressions.
```{r, warning=FALSE}
# add a row
copy_y_copy <- rbind(copy_y_copy, colSums(copy_y_copy == "1"))
copy_y_copy[nrow(copy_y_copy),]
# graph table
tissue_table <- data.frame("tissues"=colnames(copy_y_copy)[2:ncol(copy_y_copy)],
                           "num_gene"=as.numeric(copy_y_copy[nrow(copy_y_copy),2:ncol(copy_y_copy)]))

tissue_table_percent_iid <- tissue_table
tissue_table_percent_iid$per <- paste0(round(tissue_table_percent_iid$num_gene / sum(tissue_table_percent_iid$num_gene)*100, digits = 1), '%')
tissue_table_percent_iid <- tissue_table_percent_iid[-nrow(tissue_table_percent_iid),]
rownames(tissue_table_percent_iid) <- 1:nrow(tissue_table_percent_iid)

# graph
ggplot(tissue_table_percent_iid, aes(x = tissues, y = num_gene))+
  geom_text(aes(label = per), size = 2, hjust = 0.5, vjust = -0.9, position = "stack")+
  geom_col(width = 0.6) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 5.5))
```


# miR
* We conduct the same statistical analysis in miR expression table as the IID above.
```{r}
all_expr <- read.table("/Users/helending/Documents/BCB330/mir_expr_table/data/combined_all_expr_tables.csv", stringsAsFactors = FALSE)
all_expr_copy <- all_expr

tissue_names <- character()
for (i in 1:length(new_tables_0.99)){
  tissue_names[i] <- colnames(new_tables_0.99[[i]])[2]
}
all_expr_copy <- all_expr[,which(colnames(all_expr_copy) %in% tissue_names)]
all_expr <- cbind("miR" = all_expr$miR, all_expr_copy)
```

## miR statistics (number 1)
```{r}
# add a column
all_expr <- cbind(all_expr, "num_tissue" = rowSums(all_expr_copy == "1"))
# graph table
num_tissue_table <- data.frame(table(all_expr$num_tissue))
colnames(num_tissue_table) <- c("num_tissues", "num_mir")

num_tissue_table_percent <- num_tissue_table
num_tissue_table_percent$per <- paste0(round(num_tissue_table_percent$num_mir / sum(num_tissue_table_percent$num_mir)*100), '%')
rownames(num_tissue_table_percent) <- 1:nrow(num_tissue_table_percent)

# graph
ggplot(num_tissue_table_percent, aes(x = num_tissues, y = num_mir))+
  geom_text(aes(label = per), size = 3, hjust = 0.5, vjust = -1, position = "stack")+
  geom_col(width = 0.8)
```

## miR statistics (number 2)
```{r, warning=FALSE}
# add a row
all_expr <- rbind(all_expr, colSums(all_expr_copy == "1"))
tissue_table <- data.frame("tissues"=colnames(all_expr)[2:ncol(all_expr)],
                           "num_miR"=as.numeric(all_expr[nrow(all_expr),2:ncol(all_expr)]))

tissue_table_percent <- tissue_table
tissue_table_percent$per <- paste0(round(tissue_table_percent$num_miR / sum(tissue_table_percent$num_miR)*100, digits = 1), '%')
tissue_table_percent <- tissue_table_percent[-nrow(tissue_table_percent),]
rownames(tissue_table_percent) <- 1:nrow(tissue_table_percent)

# graph
ggplot(tissue_table_percent, aes(x = tissues, y = num_miR))+
  geom_text(aes(label = per), size = 3, hjust = 0.5, vjust = -0.9, position = "stack")+
  geom_col(width = 0.8) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```
## miR statistics (number 3)
Using qqplot to examine whether the data are normally distributed. 
```{r}
ggqqplot(tissue_table$num_miR)
```

# Combine IID and miR
* We compare IID gene expression and miRNA expression.
```{r}
tissue_table_percent_iid <- cbind(tissue_table_percent_iid, "type"="gene")
colnames(tissue_table_percent_iid)[2] <- "num"
tissue_table_percent <- cbind(tissue_table_percent, "type"="miR")
colnames(tissue_table_percent)[2] <- "num"


tissue_table_percent <- rbind(tissue_table_percent, tissue_table_percent_iid)
tissue_table_percent$num <- as.numeric(as.character(tissue_table_percent$num)) 

# graph
ggplot(tissue_table_percent, aes(tissues, num, fill = type)) + 
  geom_bar(position="dodge", stat="identity") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))
```

## Combine getx and miR (number 1)
```{r}
num_tissue_table_percent_gtex <- cbind(num_tissue_table_percent_gtex,"type"="gtex")
num_tissue_table_percent<- cbind(num_tissue_table_percent, "type"="miR")
colnames(num_tissue_table_percent)[2] <- "num"
colnames(num_tissue_table_percent_gtex)[2] <- "num"
num_tissue_table_percent <- rbind(num_tissue_table_percent, num_tissue_table_percent_gtex)
num_tissue_table_percent$num <- as.numeric(as.character(num_tissue_table_percent$num)) 
num_tissue_table_percent <- num_tissue_table_percent[-which(num_tissue_table_percent$num_tissues ==0),]

ggplot(num_tissue_table_percent, aes(num_tissues, num, fill = type)) + 
  geom_bar(position="dodge", stat="identity") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 10))

```

## Combine getx and miR (number 2)
```{r}
tissue_table_percent_gtex <- cbind(tissue_table_percent_gtex,"type"="gtex")
colnames(tissue_table_percent_gtex)[2] <- "num"
tissue_table_percent_gtex$tissues <- c("adipose.tissue", "heart", "kidney", "lung", "ovary", "pancreas", "prostate", "testes")
tissue_table_percent_gtex <- rbind(tissue_table_percent_gtex,
                                   c("Alzheimer.s.disease", "0", "0%", "gtex"),
                                   c("bone", "0", "0%", "gtex"),
                                   c("bone.marrow", "0", "0%", "gtex"),
                                   c("mammary.gland", "0", "0%", "gtex"),
                                   c("melanoma", "0", "0%", "gtex"),
                                   c("placenta", "0", "0%", "gtex"),
                                   c("salivary.gland", "0", "0%", "gtex"))
tissue_table_percent_gtex<- tissue_table_percent_gtex[c(1, 9, 10, 11, 2, 3, 4, 12, 13, 5, 6, 14, 7, 15, 8), ]
rownames(tissue_table_percent_gtex) <- 1:nrow(tissue_table_percent_gtex)


tissue_table_percent <- rbind(tissue_table_percent, tissue_table_percent_gtex)
tissue_table_percent$num <- as.numeric(as.character(tissue_table_percent$num)) 

ggplot(tissue_table_percent[,c(1:4)], aes(tissues, num, fill = type)) + 
  geom_bar(position="dodge", stat="identity") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 10))

```

# Tissue-specific miR-gene interaction
```{r child='/Users/helending/Documents/BCB330/mir_gene_table/final_combined.Rmd', include=TRUE, eval=TRUE}
```
# References