---
title: "R Notebook"
output: html_notebook
---

```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("/Users/helending/Documents/BCB330/code/normal_tissues/placenta"))
```
# Library
```{r}
library(knitr)
library(limma)
library(ggplot2)
```

# Download GSE28423
* Experiment: find differential expression from AA, EA prostate cancer patients
```{r, results="hide"}

URL <- "https://www.ebi.ac.uk/arrayexpress/files/E-GEOD-29100"
SDRF.file <- "E-GEOD-29100.sdrf.txt"
Data.file <- "E-GEOD-29100.raw.1.zip"
if (! file.exists(SDRF.file)){
  download.file(paste(URL,SDRF.file,sep="/"), SDRF.file)
}
if (! file.exists(Data.file)){
  download.file(paste(URL,Data.file,sep="/"), Data.file)
  unzip(Data.file)
}
```

# Select only the normal samples
* 4 samples selected
```{r}
SDRF <- read.delim(SDRF.file,check.names=FALSE,stringsAsFactors=FALSE)
norm_ind <- c(2:6)
```
# Read Agilent files
* Data Cleaning: Removed bright and dark corner positive controls
```{r, results="hide", warning=FALSE}
read_data <- limma::read.maimages(files=SDRF[norm_ind, "Array Data File"], 
                          source="agilent", 
                          green.only=TRUE, 
                          other.columns = "gIsWellAboveBG")
```
# Inspect the data
## DO NOT RUN! 
```{r}
# ====================================== Plot 1 =======================================
# only neg control
read_data <- data.frame(read_data)
# View(read_data)
neg_ctrl_ind <- which(read_data$SystematicName == "NegativeControl")
read_data <- cbind("LABEL" = 1:nrow(read_data), read_data)
read_data$LABEL[neg_ctrl_ind] <- "negControl"
neg_ctrl <- read_data[neg_ctrl_ind, ]
rows <- nrow(neg_ctrl) # 288

# the rest
rest <- read_data[-neg_ctrl_ind,]
read_data$LABEL[which(read_data$LABEL != "negControl")] <- "rest"

rows <- nrow(rest) # 13449
cols <- ncol(rest) # 10

# Build plot: summarized
# View(read_data)
boxplot(read_data[, 10] ~ read_data$LABEL, ylab="Expression", xlab="labels")
for (i in 11:ncol(read_data)){
  boxplot(read_data[, i] ~ read_data$LABEL, ylab="Expression", xlab="labels", add=TRUE)
}

# ====================================== Plot 2 =======================================

# negative control with other species

# exclude human
human <- grep(pattern="hsa", read_data$SystematicName)
length(human) # 11569
read_data <- cbind("LABEL_OTHERS"=1:nrow(read_data), read_data)
read_data$LABEL_OTHERS[human] <- "human"

# exclude positive controls
pos_control <- which(read_data$ControlType == 1)
length(pos_control) # 281

# only non-human and neg control
neg_ctrl_with_other_spec <- read_data[-c(human, pos_control), ]
read_data$LABEL_OTHERS[which(read_data$LABEL_OTHERS != "human")] <- "rest (negControl+non-human)"

boxplot(read_data[, 11] ~ read_data$LABEL_OTHERS, col=c("red", "blue"))
for (i in 12:ncol(read_data)){
  boxplot(read_data[, i] ~ read_data$LABEL_OTHERS, col=c("red", "blue"), add=TRUE)
}
```
# find human miR
```{r}
human <- grep(pattern="hsa", read_data$genes$SystematicName)
human_miR <- read_data[human,]
nrow(human_miR) # 9800
```
# Remove corners
```{r}
# Remove Bright and Dark Corners
remove <- grep(pattern="Corner", read_data$genes$SystematicName)
length(remove) # 41 brightCorner/darkCorner to be removed
read_data <- read_data[-remove, ]
nrow(read_data) # 11992
```

# Normalize by quantile method
## Background correction
```{r}
read_data_bc <- backgroundCorrect(read_data,method="minimum")
plotDensities(read_data_bc, legend=FALSE)
```
## Normalize
```{r}
read_data_norm <- normalizeBetweenArrays(read_data_bc, method="quantile")
plotDensities(read_data_norm, legend=FALSE)
```
# Pick negative control
```{r}
y_neg_ctrl <- read_data[which(read_data_norm$genes$SystematicName == "NegativeControl"),]
y_neg_ctrl <- data.frame(y_neg_ctrl)


rows <- nrow(y_neg_ctrl) # 288
cols <- ncol(y_neg_ctrl) # 13
```

# Define expression threshold
### Find the n largest values and their position
```{r, results="hide"}
# top 1% 
n <- floor(rows * (cols-8) * 0.01) # 17
y_neg_ctrl <- as.matrix(y_neg_ctrl[,9:cols])
ncol(y_neg_ctrl) # 5

## Return n largest values and position for matrix m
nlargest <- function(m, n) {
  copy_m <- m
  copy_m <- as.vector(copy_m)
  res <- order(copy_m, decreasing=TRUE)[1:n]
  pos <- arrayInd(res, dim(m), useNames = TRUE)
  list(values = m[res],position = pos)
}
# mean: 17/6=2.833333
# inspect the top 1% values
greatest_0.01 <- nlargest(y_neg_ctrl, n)
selected_0.01 <- greatest_0.01$values
table(greatest_0.01$position[,2])
#  1  2  4  5 
# 10  2  1  1 
sd(greatest_0.01$position[,2]) # 1.277446
thres <- min(greatest_0.01$values) # 79.5
```

# Define each miRNA expression
```{r}
# read in human mirna expression data\
human_miR <- data.frame(human_miR)
human_miR <- human_miR[,-8]
copy_y <- human_miR[,8:ncol(human_miR)]

# compare each expression value to thres
for(i in 1:nrow(copy_y)){
  for (j in 1:ncol(copy_y)){
    if (copy_y[i, j] >= thres){
      copy_y[i, j] <- 1
    } else{
      copy_y[i, j] <- 0
    }
  }
}

# sum the rows
copy_y <- cbind("SUM"=rowSums(copy_y), copy_y)
copy_y <- cbind("miR"=human_miR$SystematicName, copy_y)
rownames(copy_y) <- 1:nrow(copy_y)
if_ctrl_all_zero(human_miR_table=human_miR, copy_y=copy_y)
```
# save final
```{r}
unique_all <- process_human_miR(copy_y, norm_ind=norm_ind)
save_known_and_novel(unique_all_table=unique_all, tissue_name="placenta")
```

