---
title: "Retrieve tissue-specifc miRNA data"
author: "Yining Ding"
output:
  html_notebook:
    toc: yes
    toc_depth: 3
bibliography: heart.bib
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath("/Users/helending/Documents/BCB330/get_data_mir_expr/normal_tissues/heart"))
```
# Objective
The objective of this notebook is to retrieve miRNA expression data from various tissues, and processing them to extract the tissue-specific miRNA expression tables. The general workflow are 1) get data, 2) select normal or disease example accordingly, 3) select negative control group, 4) get threshold of the negative control group, 5) define human miRNA expression with that threshold, 6) finally, separate known and novel miRNA and save them into different files.

# Note
This notebook uses helper file /get_data_mir_expr/agilent_workflow_helper.R. The helper file also prints some information that can help us make some alternative decision.

# Library

The libraries used in this notebook are

  * limma [@limma]
  * ggplot2 [@ggpubr]
  * miRNAmeConverter [@miRNAmeConverter]
  
```{r, include=FALSE}
library(knitr)
library(limma)
library(ggplot2)
library(miRNAmeConverter)
```

```{r, include=FALSE}
source("/Users/helending/Documents/BCB330/get_data_mir_expr/Agilent_workflow_helper.R")
```

# Download GSE28954

I chose to download GSE28954 [@cooley2012influence] data from EBI as the arrayexpress provides a better view.
```{r, results="hide"}
URL <- "https://www.ebi.ac.uk/arrayexpress/files/E-GEOD-28954"
SDRF.file <- "E-GEOD-28954.sdrf.txt"
Data.file <- "E-GEOD-28954.raw.1.zip"
if (! file.exists(SDRF.file)){
  download.file(paste(URL,SDRF.file,sep="/"), SDRF.file)
}
if (! file.exists(Data.file)){
  download.file(paste(URL,Data.file,sep="/"), Data.file)
  unzip(Data.file)
}
```

# Select only the normal samples
* Samples selected according to the EBI array data view. We can get an overview of the data from the array express.
```{r}
SDRF <- read.delim(SDRF.file,check.names=FALSE,stringsAsFactors=FALSE)
all_sample <- SDRF[,c("Array Data File","FactorValue [DISEASE]")]

kable(head(all_sample), caption = "Overview of the data", format="html") %>%
  kable_styling(bootstrap_options = c("stripe", "hover"))

# get the ids of the selected samples
norm_ind <- which(SDRF$`FactorValue [DISEASE]` == "Healthy")
```

# Read data
```{r}
read_data <- read_data_by_limma(norm_ind)
read_data <- data.frame(read_data)
read_data <- read_data[,-c(1, 2, 3, 4, 5, 6, 7, 8, 10)]
```

# Inspect plots before picking negative control group
We plot two graph for each tissue, the first one is only picking ngeative control and the second on is picking negative control with non-human miR. We wanted to examine which plot has a greater contrast. In all the tissues, we observe that the bar plots always sugguest us to pick the negative control only for the negative control group.
```{r}
plot_a <- plot_negctrl_vs_rest(read_data)

plot_b <- plot_negctrl_plus_nonhuman_vs_rest(read_data)
```
# Extract human miR
Since we only focus on the expressivity of human miR.
```{r}
# find human miRNA
human <- grep(pattern="hsa", read_data$SystematicName)
human_miR <- read_data[human,]
```

# Data cleaning
* Find the number of human miR, corner removal report, inspect the result of the selection of the negative control group. (Whether the selection is biased or not.) The series will be discarded if the selected of the negative control group has a sample bias, i.e. selection are mostly from one sample
```{r}
neg_control <- data_cleaning(read_data)
```
# Define expression
The human miR expression is defined by the lowest expressivity value from the negative control group. We also need to make sure that all the controls have expression value of 0.
```{r, warning=FALSE}
# threshold: 
greatest_0.01 <- define_expression_threshold(y_neg_ctrl=neg_control, norm_ind=length(norm_ind))

# No bias, continue.
kable(head(table(greatest_0.01$position[,2])), caption = "Top 1% selection", format="html") %>%
  kable_styling(bootstrap_options = c("stripe", "hover"))

# define expressivity by threshold
defined <- define_miRNA_expression(human_miR_table=human_miR, thres=min(greatest_0.01$values))

# make sure that all controls have expression value = 0
if_ctrl_all_zero(human_miR_table=human_miR, copy_y=defined)
```
# Save final copies
We checked the names of the miR. The known and novel miRNA are separated and stored in different folders in __saved_expr_data/__. Final tables are all in the same format, miRNA sysmatic names, expression value, and the number of samples in this series. 
```{r}
unique_all <- process_human_miR(defined, norm_ind)
result <- save_known_and_novel(unique_all_table=unique_all, tissue_name="heart")
kable(head(result), caption = "Final saved table", format="html") %>%
  kable_styling(bootstrap_options = c("stripe", "hover"))
```
# Reference